Commutative character sheaves and geometric types for supercuspidal representations

### 2017-07-02 Geometric types.tex
Some work on the Introduction, Section 4 and Section 5.

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.}
\end{quotation}
In this paper we follow the suggestion above by showing that Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence of representations $\orho_i$ of compact $p$-adic groups $\oK^i$, for $i=0, \ldots, d$, such that $(\oK^d, \rho_d)$ is a type for a supercuspidal representation of a $p$-adic group.
Let $R$ be the ring of integers of a local field with finite residue field $k$.
The main result of \cite{Yu:models} shows how to find, for each $i=0, \ldots, d$, a smooth group scheme $\underline{G}^i$ over the ring $R$ with $\underline{G}^i(R)=\,^\circ K^i $.
The geometrization of the Yu type datum uses Lusztig's theory of character sheaves on reductive groups over finite fields, so it is necessary to assume that the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic, in order to bring his work to bear. Under this assumption we show how each representation $\orho_i$ can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, where $\cs{F}^i$ is a {\it rational virtual} sheaf complex on the Greenberg transform $G^i$ of $\underline{G}^i$, by which we mean $\cs{F}^i$ is an element in the group obtained by tensoring the Grothendieck group of such sheaves with $\QQ$. 
Writing $\trFrob{\cs{F}^i}$ for the function on $G^i(k) = \underline{G}^i(R) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the rational virtual complex $\cs{F}^i$, we show in Theorem~\ref{thm:geotypes} that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i).
\end{equation}
By this theorem, then, we obtain geometric avatars for each type in a Yu datum:
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob}}
(\underline{G}^i, \cs{F}^i).
\end{tikzcd}
\]
We refer to the pair $(\underline{G}^d, \cs{F}^d)$ as a \emph{geometric type}.

@@

Next, we broaden our scope further to encompass smooth group schemes $G$ over $\Fq$ that are not necessarily commutative.
We assume $G$ is smooth, but not that it is connected, reductive or commutative. 
The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We then define category $\CCS(G)$ for such $G$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to G\ab(k)^*$
is an isomorphism.  
Since $G\ab(k)^*$ surjects onto $G(k)^*$, it follows that for each character $\chi \in G(k)^*$ there is a commutative
character sheaf $\cs{L}$ on $G$ with $\TrFrob{G}(\cs{L}) = \chi$. Moreover, we find that pullback along the quotient $q : G \to G\ab$
defines an equivalence of categories $\CCS(G\ab) \to \CCS(G)$.  
The functor $\CCS(G) \to \CS(G)$ is not
essentially surjective, missing the kinds of linear character sheaves highlighted by Kamgarpour.

@@

The hypothesis in Theorem~\ref{thm:geotypes} -- that the geometric component group of the reductive quotient of the special fibre of the smooth group scheme $\underline{G}^0$ appearing in the Yu type datum is cyclic --  is required only because Lusztig's theory of character sheaves has the same hypothesis.
If Lusztig's theory of character sheaves can be generalized to all disconnected reductive algebraic groups, then the hypothesis in Theorem~\ref{thm:geotypes} can be removed.
\bigskip

@@

We now consider the case of a smooth group scheme without the commutativity assumption.  Our approach
is to relate linear character sheaves on $G$ to character sheaves on its abelianization $G\ab = G / G\der$.
While every character $\chi \in G(k)^*$ vanishes on $G(k)\der$, it may not vanish on $G\der(k)$.  For example,
if $k$ has odd characteristic then there are nontrivial characters $\PGL_2(k) \to \EEx$ vanishing on
$\PGL_2(k)\der = \PSL_2(k)$ (see Section \ref{sec:badchar}).  In passing to $G\ab$,
we may only hope to geometrize characters that vanish on all of $G\der(k)$.

@@

@@

\subsection{\texorpdfstring{$\SL_2(\FFF)$}{\SL(2, 3)}}

Let $G = \SL_2$ over $\FFF$.  Then $G\der = G$, but $G(k)\der$ is the subgroup of semisimple elements, isomorphic to the quaternion group $Q_8$ and of index $3$.  There is thus a nontrivial character $\chi : G(k) \to \mu_3 \subset \EEx$.  The method described above for producing a linear character sheaf corresponding to this character no longer works, since $\SL_2$ is simply connected and thus has no discrete isogenies mapping onto it.  We may still produce a local system with the correct trace of Frobenius, by pushing forward the constant sheaf along the Lang isogeny and picking out the isotypic component corresponding to $\chi$.  However, since $\SL_2$ is nonabelian, the Lang isogeny is not a homomorphism and the resulting local system is not a linear character sheaf.

Alternatively, we can use the fact that Lusztig's character sheaves provide a basis for the space of class functions \cite{lusztig:86a}*{Cor. 25.7} to give a different geometrization of $\chi$.  Lusztig's character sheaves are constructed by perverse extension of local systems on a certain stratification of $G$.  In the case of $\SL_2$, there are five strata:

@@

%\vspace{0.1in}
\noindent
Each Weil character sheaf on $\SL_2$ over $\FFF$ is the perverse extension of one of the following local systems: the rank-$1$ constant sheaf $\mathbf{1}_{C_3}$ (resp. $\mathbf{1}_{C_2}$, $\mathbf{1}_{C_2^-}$, $\mathbf{1}_{1}$,  $\mathbf{1}_{-1}$) on $C_3$ (resp. $C_2$, $C_2^-$, $1$, $-1$), or the non-trivial local system $\cs{E}_{C_2}$  (resp.  $\cs{E}_{C_2^-}$) trivialized by the  double cover of $C_2$ (resp. $C_2^-$).
The following table describes these perverse sheaves by showing how they decompose in the Grothendieck group into conjugation-equivariant standard sheaves on $\SL_2$, and also gives the values of the trace of Frobenius of each Weil character sheaf on $\SL_2$ on the $7$ conjugacy classes in $\SL_2(\FFF)$.

@@

\begin{spacing}{1.3}
\[
\begin{array}{|l|l|ccccccc|}
\hline
\text{Perverse} & \text{Standard} & \multicolumn{7}{|c|}{\text{Trace of Frobenius values}} \\ \cline{3-9}
\text{Sheaf} & \text{Sheaves} & (\begin{smallmatrix}1 & 0\\ 0 & 1\end{smallmatrix}) & (\begin{smallmatrix}2 & 0\\ 0 & 2\end{smallmatrix}) & (\begin{smallmatrix}0 & 1\\ 2 & 0\end{smallmatrix})  & (\begin{smallmatrix}1 & 1\\ 0 & 1\end{smallmatrix})  & (\begin{smallmatrix}1 & 2\\ 0 & 1\end{smallmatrix}) & (\begin{smallmatrix}2 & 1\\ 0 & 2\end{smallmatrix})  & (\begin{smallmatrix}2 & 2\\ 0 & 2\end{smallmatrix}) \\
\hline
\IC(\mathbf{1}_{C_3})  &  \mathbf{1}_{C_3}[3]  & -1 & -1 & -1 & -1 & -1 & -1 & -1 \\
 \IC(\mathbf{1}_{C_2})  &  \mathbf{1}_{C_2}[2] \oplus \mathbf{1}_1[0]  & 1 & 0 & 0 & 1 & 1 & 0 & 0 \\
 \IC(\cs{E}_{C_2})  &  \cs{E}_{C_2}[2]  & 0 & 0 & 0 &  \sqrt{-3}  &  -\sqrt{-3} & 0 & 0 \\
 \IC(\mathbf{1}_{C_2^-})  &  \mathbf{1}_{C_2^-}[2] \oplus \mathbf{1}_{-1}[0]  & 0 & 1 & 0 & 0 & 0 & 1 & 1 \\
 \IC(\cs{E}_{C_2^-})  &  \cs{E}_{C_2^-}[2]  & 0 & 0 & 0 & 0 & 0 &   \sqrt{-3}  &   -\sqrt{-3}  \\
 \IC(\mathbf{1}_1)  &  \mathbf{1}_1[0]  & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
 \IC(\mathbf{1}_{-1})  &  \mathbf{1}_{-1}[0]  & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
\hline
\end{array}
\]
\end{spacing}

\vspace{0.1in}
\noindent Using the trace of Frobenius values, we find that the rational virtural Weil sheaf 
\[
- \IC(\mathbf{1}_{C_3}) - \frac32 \IC(\mathbf{1}_{C_2}) + \frac12 \IC(\cs{E}_{C_2}) - \frac32 \IC(\mathbf{1}_{C_2^-}) - \frac12 \IC(\cs{E}_{C_2^-}) + \frac32 \IC(\mathbf{1}_1) + \frac32 \IC(\mathbf{1}_{-1})
\]
also geometrizes $\chi$.  Note that all seven character sheaves appear in this geometrization, even though we start just with a character.

@@

@@

Kamgarpour gives an example \cite{kamgarpour:09a}*{(1.1)} of a linear character sheaf
on $G = \PGL_n$ that does not vanish on $G\der = \PGL_n$, defined by the extension
\[
1 \to \mu_n \to \SL_n \to \PGL_n \to 1
\]
together with a non-trivial character of $\mu_n(k)$. \todo{Small change here.}
In order to obtain a relationship between linear character sheaves on $G$ and characters of $G(k)$,
he opts to give a different definition of commutator and, in doing so, introduces a `stacky abelianization' of $G$ in order to geometrize characters of $G(\Fq)$.
 Since we have already seen the need to adapt the
notion of character sheaf, even in the commutative case, we instead add restrictions
to the definition of commutative character sheaf and leave the definition of $G\der$ unchanged, allowing us to use the schematic abelianization of $G$ in the geometrization of characters of $G(\Fq)$.

@@

\subsection{Quasicharacters of smooth group schemes over certain henselian traits}

Recall that $R$ is the ring of integers of a local field with finite residue field $k$.
The maximal idea of $R$ will be denoted by $\mathfrak{m}$. 
Let $\underline{G}$ be a smooth group scheme over $R$.
Here we shall use  \cite{bertapelle-gonzales:Greenberg} for the definition and fundamental properties of the Greenberg transform.
Let $G$ be the Greenberg transform of $\underline{G}$; then $G$ is a group scheme over $\Fq$ and there is a canonical isomorphism
\[
G(\Fq) = \underline{G}(R).
\]

\begin{proposition}\label{prop:quasicharacters}
For every quasicharacter character $\varphi : \underline{G}(R) \to \EEx$ there is a Weil sheaf $\cs{L}$ on $G$ such that \[\trFrob{\cs{L}} =  \varphi.\] \end{proposition}

\begin{proof}
By continuity of $\varphi : \underline{G}(R) \to \EE^\times$, there is some $m \in \NN$ and a factorization
\[
\begin{tikzcd}
\underline{G}(R) \arrow{rr}{\varphi} \arrow{rd} && \EEx\\
& \underline{G}(R/\mathfrak{p}^{m+1}) \arrow{ru}[swap]{\varphi_m} 
\end{tikzcd}
\] 
Set $R_m = R/\mathfrak{p}^{m+1}$ and
set $G_m = \Gr_m^{R}(\underline{G})$, the Greenberg transform of $\underline{G}\times_{\Spec{R}}\Spec{R_m}$.
Then $G_m$ is a smooth group scheme over $\Fq$ and $G_m(\Fq) = \underline{G}(R_m)$.
Using Theorem~\ref{thm:geo}, let $\cs{L}_m$ be a geometrization of the character $\varphi_m: G_m(\Fq) \to \EEx$; so
\[
\trFrob{\cs{L}_m} = \varphi_m
\]
on $G_m(\Fq)$.
%
Recall that the full Greenberg transform $G \ceq \Gr^{R}(\underline{G})$ is a group scheme over $\Fq$ such that $G(\Fq) = \underline{G}(R)$; it comes equipped with a morphism $G \to G_m$.
Let $\cs{L}$ be the Weil sheaf on $G$ obtained from $\cs{L}_m$ by pullback along $G \to G_m$. 
Then $\cs{L}$ is a quasicharacter sheaf on $G$, in the sense of \cite{cunningham-roe:13a}*{Def 4.2}, such that $\trFrob{\cs{L}} = \varphi$.
\end{proof}

@@

There is a Weil sheaf complex $\cs{K}_\psi$ on $\Sp(V)\ltimes V^\sharp$ \cite{gurevich-hadani:07a}*{Theorem 3.2.2.1} (see also \cite{gurevich-hadani:11a}*{Theorem 4.5}) such that 
\begin{equation}
\trFrob{\cs{K}_\psi} = \trace(W_\psi).
\end{equation}
%
Since $\cs{K}_\psi$ is an object in Deligne's category $D^b_c(\Sp(V)\ltimes V^\sharp,\EE)$, the left hand side of this equality must be interpreted accordingly.

@@

\subsection{Geometrization of characters of types}\label{ssec:geotypes}

Finally, we come to the main result of Section~\ref{sec:types}.
%
Since Yu's theory refers to complex representations, and since our geometrization uses $\ell$-adic sheaves, we grit our teeth and fix an isomorphism $\CC \approx \EE$.

>>>>

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.}
\end{quotation}
In this paper we follow the suggestion above by showing that certain Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence of representations $\orho_i$ of compact $p$-adic groups $\oK^i$, for $i=0, \ldots, d$, such that $(\oK^d, \rho_d)$ is a type for a supercuspidal representation of a $p$-adic group.
Let $R$ be the ring of integers of a local field with finite residue field $k$.
The main result of \cite{Yu:models} shows how to find, for each $i=0, \ldots, d$, a smooth group scheme $\underline{G}^i$ over the ring $R$ with $\underline{G}^i(R)=\,^\circ K^i $.
The geometrization of the Yu type datum that we consider in this paper uses Lusztig's theory of character sheaves on reductive groups over finite fields, for which it is necessary to assume that the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic. 
We must also place a condition on the linear characters that appear in the Yu datum, 
Under this assumption we show how each representation $\orho_i$ can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, where $\cs{F}^i$ is a {\it rational virtual} sheaf complex on the Greenberg transform $G^i$ of $\underline{G}^i$, by which we mean $\cs{F}^i$ is an element in the group obtained by tensoring the Grothendieck group of such sheaves with $\QQ$. 
Writing $\trFrob{\cs{F}^i}$ for the function on $G^i(k) = \underline{G}^i(R) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the rational virtual complex $\cs{F}^i$, we show in Theorem~\ref{thm:geotypes} that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i).
\end{equation}
By this theorem, then, we obtain geometric avatars for each type in a Yu datum:
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob}}
(\underline{G}^i, \cs{F}^i).
\end{tikzcd}
\]
We refer to the pair $(\underline{G}^d, \cs{F}^d)$ as a \emph{geometric type}.

@@

Next, we broaden our scope further to encompass smooth group schemes $G$ over $\Fq$ that are not necessarily commutative.
We assume $G$ is smooth, but not that it is connected, reductive or commutative. 
The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We then define category $\CCS(G)$ for such $G$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to G\ab(k)^*$
is an isomorphism.  
Since $G\ab(k)^*$ surjects onto $G(k)^*$, it follows that for each character $\chi \in G(k)^*$ trivial on $G\der(k)$ there is a commutative
character sheaf $\cs{L}$ on $G$ with $\TrFrob{G}(\cs{L}) = \chi$. Moreover, we find that pullback along the quotient $q : G \to G\ab$
defines an equivalence of categories $\CCS(G\ab) \to \CCS(G)$.  
The functor $\CCS(G) \to \CS(G)$ is not
essentially surjective, missing the kinds of linear character sheaves highlighted by Kamgarpour.

@@

The hypothesis in Theorem~\ref{thm:geotypes} -- that the geometric component group of the reductive quotient of the special fibre of the smooth group scheme $\underline{G}^0$ appearing in the Yu type datum is cyclic --  is required only because Lusztig's theory of character sheaves has the same hypothesis.
If Lusztig's theory of character sheaves can be generalized to all disconnected reductive algebraic groups, then the hypothesis in Theorem~\ref{thm:geotypes} can be removed.
%
The second hypothesis  in Theorem~\ref{thm:geotypes} -- that the quasicharacters $\varphi^i$ appearing in the Yu type datum satisfy the hypothesis of Proposition~\ref{prop:quasicharacters} -- ensures that we may use Section~\ref{sec:noncom} to find a geometric type which equivariant for conjugation, as defined in Section~\ref{ssec:equivariant}.
Section~\ref{sec:badchar} presents some evidence to suggest that this hypothesis is not always necessary. 
\bigskip

@@

We now consider the case of a smooth group scheme without the commutativity assumption.  Our approach
is to relate linear character sheaves on $G$ to character sheaves on its abelianization $G\ab = G / G\der$, where $G\der$ is defined by \cite{demazure:SGA3-VIB}*{D\'efinition 7.2.2} and the quotient is $G\ab$ is an instance of  \cite{demazure:SGA3-VIA}*{Section 7.2.2}.
While every character $\chi \in G(k)^*$ vanishes on $G(k)\der$, it may not vanish on $G\der(k)$.  For example,
if $k$ has odd characteristic then there are nontrivial characters $\PGL_2(k) \to \EEx$ vanishing on
$\PGL_2(k)\der = \PSL_2(k)$ (see Section \ref{sec:badchar}).  In passing to $G\ab$,
we may only hope to geometrize characters that vanish on all of $G\der(k)$.

@@

\subsection{Equivariance Weil sheaf complexes (move to somewhere else?)}\label{ssec:equivariant}

Suppose $(\cs{L},\mu)$ is a linear character sheaf on $G$. 
As the diagram below explains, the isomorphism $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ determines an isomorphism $a^*\cs{L} \to p_2^*\cs{L}$ of Weil local systems on $G\times G$, where  $a : G\times G\to G$ is conjugation $a(x,y)=xyx^{-1}$ and $p_2: G\times G\to G$ is projection $p_2(x,y)=y$.
\[
\begin{tikzcd}[column sep = 30]
a^*\cs{L} \arrow[equal]{d}[swap]{a= m\circ (m\times (i\circ p_1))} \arrow[dotted]{r} & p_2^*\cs{L}  \\
(m\times (i\circ p_1))^* m^*\cs{L} \arrow{d}[swap]{(m\times (i\circ p_1))^*\mu}  & p_1^*(\cs{L} \otimes \cs{L}^\vee) \otimes p_2^*\cs{L} \arrow[equal]{u}[swap]{\cs{L}\otimes \cs{L}^\vee \iso (\EE)_{G}}\\
m^*\cs{L}\otimes (i\circ p_1))^*\cs{L} \arrow{r}{\mu \otimes \id} & p_1^*\cs{L} \otimes p_2^*\cs{L} \otimes p_1^*\cs{L}^\vee \arrow[equal]{u}
\end{tikzcd}
\]
Although every linear character sheaf on $G$ is a conjugation-equivariant Weil local systems on $G$, the converse is certainly not true.

For connected reductive groups $G$ over $k$, Lusztig's character sheaves are simple perverse sheaves $\IC(C,\mathcal{L})$ on $G$ that come equipped with an isomorphism $a^*\, \IC(C,\mathcal{L}) \to p_2^*\, \IC(C,\mathcal{L})$ in Deligne's category $D^b_c(G;\EE)$.
Consequently, when $\IC(C,\mathcal{L})$ is also a Weil sheaf complex, its trace of Frobenius function $\trFrob{\IC(C,\cs{L})} : G(k)\to \EE$ is a class function, called the characteristic function of $\IC(C,\cs{L})$.
By \cite{lusztig:86a}*{Cor. 25.7}, Weil character sheaves on $G$ determine a basis for the $\EE$-space of class functions on $G(k)$.
In particular, any character of any representation of $G(k)$ can be expressed as a ${\bar \QQ}$-linear combination of characteristic functions of Weil character sheaves on $G$.
We will refer to a formal sum $\sum_{i\in I} a_i \cs{F}_i$ of conjugation-equivariant Weil $\ell$-adic sheaf complexes with ${\bar \QQ}$-coefficients as a rational virtual equivariant Weil sheaf complex on $G$.
When specialized to the case of characters $\chi$ of $G(k)$, Lusztig's theory of character sheaves geometrizes $\chi$ as a rational virtual equivariant Weil sheaf complex on $G$.
We see an example of this in Section~\ref{ssec:SL23}.

@@

\subsection{\texorpdfstring{$\SL_2(\FFF)$}{\SL(2, 3)}}\label{ssec:SL23}

Let $G = \SL_2$ over $\FFF$.  Then $G\der = G$, but $G(k)\der$ is the subgroup of semisimple elements, isomorphic to the quaternion group $Q_8$ and of index $3$.  There is thus a nontrivial character $\chi : G(k) \to \mu_3 \subset \EEx$.  The method described above for producing a linear character sheaf corresponding to this character no longer works, since $\SL_2$ is simply connected and thus has no discrete isogenies mapping onto it.  We may still produce a local system with the correct trace of Frobenius, by pushing forward the constant sheaf along the Lang isogeny and picking out the isotypic component corresponding to $\chi$.\todo{So, maybe not?}  However, since $\SL_2$ is nonabelian, the Lang isogeny is not a homomorphism and the resulting local system is not a linear character sheaf.

Alternatively, we can use the fact that Lusztig's character sheaves provide a basis for the space of class functions \cite{lusztig:86a}*{Cor. 25.7} to give a different geometrization of $\chi$, as discussed in Section~\ref{ssec:equivariant}.  Lusztig's character sheaves are constructed by perverse extension of local systems on a certain stratification of $G$.  In the case of $\SL_2$, there are five strata:

@@

@@

\noindent
Each Weil character sheaf on $\SL_2$ over $\FFF$ is the perverse extension of one of the following local systems: the rank-$1$ constant sheaf $\mathbf{1}_{C_3}$ (resp. $\mathbf{1}_{C_2}$, $\mathbf{1}_{C_2^-}$, $\mathbf{1}_{1}$,  $\mathbf{1}_{-1}$) on $C_3$ (resp. $C_2$, $C_2^-$, $1$, $-1$), or the non-trivial rank-$1$ local system $\cs{E}_{C_2}$  (resp.  $\cs{E}_{C_2^-}$) on $C_2^{-1}$  (resp.  $\cs{E}_{C_2^-}$) trivialized by the double cover of $C_2$ (resp. $C_2^-$).
The following table describes these perverse sheaves by showing how they decompose in the Grothendieck group into standard sheaves on $\SL_2$, and also gives the values of the trace of Frobenius on the $7$ conjugacy classes in $\SL_2(\FFF)$.

\begin{spacing}{1.3}
\[
\begin{array}{|l|l|ccccccc|}
\hline
\text{Perverse} & \text{Standard} & \multicolumn{7}{|c|}{\text{Trace of Frobenius values at conjugacy classes}} \\ \cline{3-9}
\text{Sheaf} & \text{Sheaves} & (\begin{smallmatrix}1 & 0\\ 0 & 1\end{smallmatrix}) & (\begin{smallmatrix}2 & 0\\ 0 & 2\end{smallmatrix}) & (\begin{smallmatrix}0 & 1\\ 2 & 0\end{smallmatrix})  & (\begin{smallmatrix}1 & 1\\ 0 & 1\end{smallmatrix})  & (\begin{smallmatrix}1 & 2\\ 0 & 1\end{smallmatrix}) & (\begin{smallmatrix}2 & 1\\ 0 & 2\end{smallmatrix})  & (\begin{smallmatrix}2 & 2\\ 0 & 2\end{smallmatrix}) \\
\hline
\IC(\mathbf{1}_{C_3})  &  \mathbf{1}_{C_3}[3]  & -1 & -1 & -1 & -1 & -1 & -1 & -1 \\
 \IC(\mathbf{1}_{C_2})  &  \mathbf{1}_{C_2}[2] \oplus \mathbf{1}_1[0]  & 1 & 0 & 0 & 1 & 1 & 0 & 0 \\
 \IC(\cs{E}_{C_2})  &  \cs{E}_{C_2}[2]  & 0 & 0 & 0 &  \sqrt{-3}  &  -\sqrt{-3} & 0 & 0 \\
 \IC(\mathbf{1}_{C_2^-})  &  \mathbf{1}_{C_2^-}[2] \oplus \mathbf{1}_{-1}[0]  & 0 & 1 & 0 & 0 & 0 & 1 & 1 \\
 \IC(\cs{E}_{C_2^-})  &  \cs{E}_{C_2^-}[2]  & 0 & 0 & 0 & 0 & 0 &   \sqrt{-3}  &   -\sqrt{-3}  \\
 \IC(\mathbf{1}_1)  &  \mathbf{1}_1[0]  & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
 \IC(\mathbf{1}_{-1})  &  \mathbf{1}_{-1}[0]  & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
\hline
\end{array}
\]
\end{spacing}

\vspace{0.1in}
\noindent Using the trace of Frobenius values, we find that the rational virtual equivariant Weil sheaf complex
\[
- \IC(\mathbf{1}_{C_3}) - \frac32 \IC(\mathbf{1}_{C_2}) + \frac12 \IC(\cs{E}_{C_2}) - \frac32 \IC(\mathbf{1}_{C_2^-}) - \frac12 \IC(\cs{E}_{C_2^-}) + \frac32 \IC(\mathbf{1}_1) + \frac32 \IC(\mathbf{1}_{-1})
\]
also geometrizes $\chi$.  Note that all seven character sheaves appear in this geometrization, even though we start just with a character.

@@

\todo{}
Explain that while $(f_! \EE)_\psi$ is not a commutative character sheaf, it does satisfy a weaker version of Definition~\ref{def:CCS}. Namely, there is an isomorphism $\gamma$ such that
 \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d}[swap]{\gamma} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} &  \theta^*(\cs{L}\boxtimes\cs{L}).
  \end{tikzcd}
  \]

@@

Kamgarpour gives an example \cite{kamgarpour:09a}*{(1.1)} of a linear character sheaf
on $G = \PGL_n$ that does not vanish on $G\der = \PGL_n$, defined by the extension
\[
1 \to \mu_n \to \SL_n \to \PGL_n \to 1
\]
together with a non-trivial character of $\mu_n(k)$. 
In order to obtain a relationship between linear character sheaves on $G$ and characters of $G(k)$,
he opts to give a different definition of commutator and, in doing so, introduces a `stacky abelianization' of $G$ in order to geometrize characters of $G(\Fq)$.
 Since we have already seen the need to adapt the
notion of character sheaf, even in the commutative case, we instead add restrictions
to the definition of commutative character sheaf and leave the definition of $G\der$ unchanged, allowing us to use the schematic abelianization of $G$ in the geometrization of characters of $G(\Fq)$.

@@

\subsection{Quasicharacters of smooth group schemes over certain henselian traits}\label{ssec:quasicharacters}

Recall that $R$ is the ring of integers of a local field with finite residue field $k$.
The maximal idea of $R$ will be denoted by $\mathfrak{m}$. 
Let $\underline{G}$ be a smooth group scheme over $R$.
Here we shall use  \cite{bertapelle-gonzales:Greenberg} for the definition and fundamental properties of the Greenberg transform.
Let $G$ be the Greenberg transform of $\underline{G}$; then $G$ is a group scheme over $\Fq$ and there is a canonical isomorphism
\[
G(\Fq) = \underline{G}(R).
\]
Let $\varphi : \underline{G}(R) \to \EEx$ be a quasicharacter. 
By the continuity of $\varphi$ there is some $r \in \NN$ and a factorization
\[
\begin{tikzcd}
\underline{G}(R) \arrow{rr}{\varphi} \arrow{rd} && \EEx\\
& \underline{G}(R/\mathfrak{p}^{r+1}) \arrow{ru}[swap]{\varphi_r} 
\end{tikzcd}
\] 
The least such $r\in \NN$ is called the {\it level} of $\varphi$.
For any $r\in \NN$, define $R_r \ceq R/\mathfrak{p}^{r+1}$ and
set $G_r \ceq \Gr_r^{R}(\underline{G})$, the Greenberg transform of $\underline{G}\times_{\Spec{R}}\Spec{R_r}$.
Then $G_r$ is a smooth group scheme over $\Fq$ and $G_r(\Fq) = \underline{G}(R_r)$.

\begin{proposition}\label{prop:quasicharacters}
Let $\underline{G}$ be a smooth group scheme over $R$.
Let $\varphi : \underline{G}(R) \to \EEx$ be quasicharacter and let $r$ be the level of $\varphi$.
If $\varphi_r : G_m(k)\to \EEx$ is trivial on $G_{r,\operatorname{der}}(k)$ then 
there is a commutative character sheaf $\cs{L}_r\in \CCS(G_r)$ such that 
\[
\trFrob{\cs{L}_r} =  \varphi_r.
\] 
\end{proposition}

\begin{proof}
This is a direct consequence of Theorem~\ref{thm:geo}.
%let $\cs{L}_r$ be a geometrization of the character $\varphi_r: G_r(\Fq) \to \EEx$.
\end{proof}

With $\varphi$ as above, note that if $r'$ is any integer greater or equal to the level $r$ of $\varphi$, then the pullback of $\cs{L}_r$ along $G_{r'} \to G_{r}$ is a linear character sheaf on $G_r$. 
Thus, $\cs{L}_{r'}$ is conjugation-equivariant in the sense of Section~\ref{ssec:equivariant}

$a_r : G_r\times G_r \to G_r$ is conjugation for $G_r$, then this determines an isomorphism $a_r^*\cs{L}_r \iso (\EE)_{G_r}\boxtimes \cs{L}_r$, so $\cs{L}_r$ is

Recall that the full Greenberg transform $G \ceq \Gr^{R}(\underline{G})$ is a group scheme over $\Fq$ such that $G(\Fq) = \underline{G}(R)$; it comes equipped with a morphism $G \to G_m$.
The Weil sheaf on $G$ obtained from $\cs{L}_m$ by pullback along the morphism of group schemes $G \to G_m$ is a quasicharacter sheaf on $G$, in the sense of \cite{cunningham-roe:13a}*{Def 4.2}, such that 
\[
\trFrob{\cs{L}} = \varphi.
\]

@@

By \cite{gurevich-hadani:07a}*{Theorem 3.2.2.1} (see also \cite{gurevich-hadani:11a}*{Theorem 4.5}), there is a Frobenius-stable conjugation-equivariant perverse sheaf $\cs{K}_\psi$ on $\Sp(V)\ltimes V^\sharp$  such that 
\begin{equation}\label{eqn:Jacobi}
\trFrob{\cs{K}_\psi} = \trace(W_\psi).
\end{equation}
%
Since $\cs{K}_\psi$ is an object in Deligne's category $D^b_c(\Sp(V)\ltimes V^\sharp,\EE)$, the left hand side of this equality must be interpreted accordingly.

@@

\subsection{Geometrization of characters of certain types}\label{ssec:geotypes}

We may now give the main result of Section~\ref{sec:types}.
%
Since Yu's theory refers to complex representations, and since our geometrization uses $\ell$-adic sheaves, we grit our teeth and fix an isomorphism $\CC \approx \EE$.
From Section~\ref{ssec:equivariant}, recall the definition of a rational virtual equivariant Weil sheaf complex.

\begin{theorem}\label{thm:geotypes}
Let $\oK^i$ $\orho^0$, $\varphi^i$, for $i=0, \ldots d$, be a Yu type datum as in Section~\ref{ssec:review}, \ref{Y0}, \ref{Y1} and \ref{Y2}.
Let $\underline{G}^i$ and $\varrho_0\red$ be as in Section~\ref{ssec:review}, \ref{M0} and \ref{M1}.
Assume $\pi_0((\underline{G}^0)_\Fq^\reductive)$ is cyclic.
For $i=0, \ldots ,d$, let $G^i = \Gr_{R}(\underline{G}^i)$ be the Greenberg transform of the smooth group scheme $\underline{G}^i$ appearing in Section~\ref{ssec:review}.
Suppose that the quasicharacters $\varphi^i$, given in Section~\ref{ssec:review}, \ref{M2}, satisfy the hypothesis of Proposition~\ref{prop:quasicharacters}.
Then there is a rational virtual equivariant Weil sheaf complex $\cs{F}_i$ on $G^i$ such that $\trFrob{\cs{F}_i} = \trace(\orho_i)$, for $i=0, \ldots , d$.
\end{theorem}
